<small> Supplementary materials for the main submission entitled - <br><br> Clustering time series based on probability distributions across temporal granularities

### 2021-11-09 paper-supplementary.Rmd
code ready but document not running

```{r external, include = FALSE, cache = FALSE}
knitr::read_chunk('script/main.R')
```

@@

@@

(ref:missing-data) Investigating the temporal location of missing values for customers who have implicit missing values. There are $13,735$ customers in the data set, with 8,685 having no missing values and the remaining $5,050$ having at least one missing value. Each cross represents a missed observation in time, while the line connecting two dots represents continuous missingness over time. Missing values occur at random times and do not appear to follow a pattern, although there is a higher concentration of missing values in September and October 2012 for the majority of customers. This plot is inspired by @wang2020tsibble.

@@

_Prototype selection method_

@@

```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud."}
knitr::include_graphics("figs/liminal_view.png")
```

@@

<!-- ### Interaction of granularities -->

<!-- Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$. In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories. Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs. Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively. D4 varies across both $g1$ and $g2$ categories. D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same. -->

@@

>>>>

```{r external, include = FALSE, cache = FALSE}
knitr::read_chunk('script/main.R')
read_chunk(here::here("validation/simdata01.R"))
```

@@

# Raw data plot

@@

# Missing data plot
(ref:missing-data) Investigating the temporal location of missing values for customers who have implicit missing values. There are $13,735$ customers in the data set, with 8,685 having no missing values and the remaining $5,050$ having at least one missing value. Each cross represents a missed observation in time, while the line connecting two dots represents continuous missingness over time. Missing values occur at random times and do not appear to follow a pattern, although there is a higher concentration of missing values in September and October 2012 for the majority of customers. This plot is inspired by @wang2020tsibble.

@@

# Prototype selection

@@

```{r mds, out.width="100%", fig.cap="Instance selection using tours and projecting the points in a lower dimensional tsne cloud., eval = FALSE"}
knitr::include_graphics("figs/liminal_view.png")
```

@@

# Interaction of granularities

Consider a case in which there are only two interacting granularities of interest, $g1$ and $g2$. In contrast to the previous situation, when we could study distributions across $n_{g_1} + n_{g_2} = 5$ separate categories, with interaction, we must evaluate the distribution of the $n_{g_1}*n_{g_2}=6$ combination of categories. Consider the $4$ designs in Figure \ref{fig:interaction-gran}, where various distributions are assumed for different combinations of categories, resulting in different designs. Design $D1$ exhibits no change in distributions across $g1$ or $g2$, whereas Designs $D2$ and $D3$ alter across only $g1$ and $g2$, respectively. D4 varies across both $g1$ and $g2$ categories. D3 and D4 appear similar based on their relative differences across consecutive categories, but $D4$ also changes across facets, unlike $D3$, which has all facets look the same.

@@

```{r sim-val}
set.seed(9999)
nx_val = 2 # number of x-axis levels
nfacet_val = 3 # number of facet levels
w1_val = 3 # increment in mean
w2_val = 0 # increment in sd
mean_val = 0 # mean of normal distribution of starting combination
sd_val = 1 # sd of normal distribution of starting combination
quantile_prob_val = seq(0.1, 0.9, 0.1)
```

```{r nobs50}
ntimes_val = 300 # nobs per combination 
```

<!-- # Clustering on simulated datasets -->

```{r sample-seed-few}
sample_seed <-  seq(10, 100, 10)
```

<!-- _DGP: Generate 10 time series from each designs_  -->
<!-- Time series are simulated from each of these designs with $50$ observations in each group. So we have $50$ observations each for the six combination of categories $(1,1), (1, 2), (2, 1), (2, 2), (3, 1), (3, 2)$. Time series are simulated for ten different seeds (from `sample_seed`) for each design. `data_varall-1` represents a data set from a design $D_{var_{all}}$ (distributions change across both facet and x) with a seed $10$. `data_varall-2` represents a data set from a design $D_{var_{all}}$ with a seed $20$ and so on. -->
<!-- `data_null`, `data_varx` and `data_varf `corresponds to designs $D_{null}$, $D_{var_{x}}$ and $D_{var_{f}}$ designs respectively.    -->

```{r change-index-data-few}
change_index_data <- function(ntimes_val = NULL,
                         nx_val = NULL,
                         nfacet_val = NULL,
                         sim_function = sim_varx_normal){

  data <- sim_panel(
    nx = nx_val, nfacet =  nfacet_val,
    ntimes = ntimes_val,
    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
  ) %>% unnest(data)

  index_new <- map(seq_len(ntimes_val), function(i){
    map((seq_len(nx_val*nfacet_val)), function(j)
    {
      value = i + (j-1)*ntimes_val
    })
  }) %>% unlist()

  data_new = data %>%
    ungroup() %>%
    mutate(index_old = row_number(),
           index_new = index_new)

  y = data_new[match(index_new, data_new$index_old),]

  y <- y %>%
    mutate(time = row_number()) %>%
    select(-c(index_old, index_new))

  return(y)
}
```

```{r data-make-few}
data_null <- map(sample_seed, function(seed){
  set.seed(seed)
  change_index_data(5, 2, 3, sim_null_normal) %>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_null")

data_varf <- map(sample_seed, function(seed){
   set.seed(seed)
  change_index_data(5, 2, 3, sim_varf_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varf")

data_varx <- map(sample_seed, function(seed){
   set.seed(seed)
  change_index_data(5, 2, 3, sim_varx_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varx")

data_varall <- map(sample_seed, function(seed){
   set.seed(seed)
  change_index_data(5, 2, 3, sim_varall_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varall")

data_q <- bind_rows(data_null,
                      data_varf,
                      data_varx,
                      data_varall) %>%
  mutate(seed_id = sprintf("%02d", 
                              as.numeric(seed_id))) %>% 
  mutate(unique_data = paste(data_type, seed_id, sep = "-"))

data_q$data_type <- factor(data_q$data_type,
                                    levels = c("data_null" , "data_varf",
                                               "data_varx","data_varall"))

```{r JS-few}

```{r JS-mydata-few}
data_q_wide <- data_q %>%
  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
  pivot_wider(names_from = unique_data,
              values_from = sim_data_quantile) %>%
  select(-c(1, 2))

ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
ldata <- nrow(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist) %>%
  mutate(data_type1 = as.numeric(data_type1)) %>%
  left_join(ndata, by = c("data_type1" = "index"))

dist_mat_format <- dist_mat %>%
  ungroup() %>%
  select(-data_type1, -unique_data)

rownames(dist_mat_format) <- dist_mat$unique_data
```

```{r hier-clust-few}
d = stats::as.dist(dist_mat_format)
par(cex=0.5)
hc = stats::hclust(d,method="complete")
plot(hc)
```

```{r mds-plot-interact-few}
mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- dist_mat$unique_data

groups<-cutree(hc, k=4)

all_data_cluster <- cbind(dist_mat_format, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups)) %>% as_tibble()

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups, label = rownames(mds))) +
  geom_point(size = 1) +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2") +
  coord_fixed(ratio = 1)+
  theme(text = element_text(size = 6))

mds_plot_10
```

<!-- # Clustering on simulated datasets -->

```{r sample-seed}
sample_seed <-  seq(1, 100, 1)
```

<!-- _DGP: Generate 10 time series from each designs_  -->
<!-- Time series are simulated from each of these designs with $50$ observations in each group. So we have $50$ observations each for the six combination of categories $(1,1), (1, 2), (2, 1), (2, 2), (3, 1), (3, 2)$. Time series are simulated for ten different seeds (from `sample_seed`) for each design. `data_varall-1` represents a data set from a design $D_{var_{all}}$ (distributions change across both facet and x) with a seed $10$. `data_varall-2` represents a data set from a design $D_{var_{all}}$ with a seed $20$ and so on. -->
<!-- `data_null`, `data_varx` and `data_varf `corresponds to designs $D_{null}$, $D_{var_{x}}$ and $D_{var_{f}}$ designs respectively.    -->

```{r change-index-data}
change_index_data <- function(ntimes_val = NULL,
                         nx_val = NULL,
                         nfacet_val = NULL,
                         sim_function = sim_varx_normal){

  data <- sim_panel(
    nx = nx_val, nfacet =  nfacet_val,
    ntimes = ntimes_val,
    # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
    sim_dist = sim_function(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
  ) %>% unnest(data)

  index_new <- map(seq_len(ntimes_val), function(i){
    map((seq_len(nx_val*nfacet_val)), function(j)
    {
      value = i + (j-1)*ntimes_val
    })
  }) %>% unlist()

  data_new = data %>%
    ungroup() %>%
    mutate(index_old = row_number(),
           index_new = index_new)

  y = data_new[match(index_new, data_new$index_old),]

  y <- y %>%
    mutate(time = row_number()) %>%
    select(-c(index_old, index_new))

  return(y)
}
```

```{r data-make}
data_null <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_null_normal) %>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_null")

data_varf <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varf_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varf")

data_varx <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varx_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varx")

data_varall <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varall_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varall")

data_q <- bind_rows(data_null,
                      data_varf,
                      data_varx,
                      data_varall) %>%
  mutate(seed_id = sprintf("%02d", 
                              as.numeric(seed_id))) %>% 
  mutate(unique_data = paste(data_type, seed_id, sep = "-"))

data_q$data_type <- factor(data_q$data_type,
                                    levels = c("data_null" , "data_varf",
                                               "data_varx","data_varall"))

```{r JS}

```{r JS-mydata}
data_q_wide <- data_q %>%
  select(id_facet, id_x, unique_data, sim_data_quantile) %>%
  pivot_wider(names_from = unique_data,
              values_from = sim_data_quantile) %>%
  select(-c(1, 2))

ndata <- data_q %>% distinct(unique_data) %>% mutate(index = row_number())
ldata <- nrow(ndata)
lcomb <-  nx_val*nfacet_val

dist_data <- map(1:ldata, function(x){ # first data
  map(1:ldata, function(y){ # 2nd data
    map(1:lcomb, function(z){ # number of combinations nx*nfacet
      JS(
        prob = quantile_prob_val,
        unlist(data_q_wide[z,x]),
        unlist(data_q_wide[z,y])
      ) %>% as_tibble()
    })%>% bind_rows(.id = "combinations")
  })%>% bind_rows(.id = "data_type1")
}) %>% bind_rows(.id = "data_type2")

dist_mat <- dist_data %>%
  group_by(data_type1, data_type2) %>%
  summarise(dist = sum(value)) %>%
  pivot_wider(names_from = data_type2,
              values_from = dist) %>%
  mutate(data_type1 = as.numeric(data_type1)) %>%
  left_join(ndata, by = c("data_type1" = "index"))

dist_mat_format <- dist_mat %>%
  ungroup() %>%
  select(-data_type1, -unique_data)

rownames(dist_mat_format) <- dist_mat$unique_data
```

```{r hier-clust}
d = stats::as.dist(dist_mat_format)
par(cex=0.3)
hc = stats::hclust(d,method="complete")
plot(hc)
```

```{r mds-plot-interact}
mds <- d %>%
  cmdscale() %>%
  as_tibble()

colnames(mds) <- c("Dim.1", "Dim.2")
rownames(mds) <- dist_mat$unique_data

groups<-cutree(hc, k=4)

all_data_cluster <- cbind(dist_mat_format, groups) %>%
  cbind(mds)%>%
  mutate(groups = as.factor(groups)) %>% as_tibble()

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups,
       label = rownames(mds))) +
  geom_point(size = 0.5) +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2") +
  coord_fixed(ratio = 1)

mds_plot_10
```

```{r heatmap-interact, fig.height=8}
heatmap_raw <- data_q %>% 
  #select(unique_data, sim_data_quantile, id_facet, id_x) %>% 
  mutate(category = paste0("facet= ", id_facet, ",x= ", id_x)) %>% 
  unnest(sim_data_quantile) %>% 
  group_by(unique_data, category) %>% 
  mutate(D = row_number()) %>% 
  arrange(D, category) %>% 
  left_join(dist_mat, by = c("unique_data")) %>% 
  select(c(2:9))

data2 <- cbind(dist_mat_format, groups)

heatmap_final <- bind_cols(serial_data = rownames(data2) %>%
                             as_tibble(),
          group = data2$groups) %>%
  left_join(heatmap_raw, by = c("value" = "unique_data")) %>% select(-id_facet, - id_x, - data_type1) %>% 
  arrange(D, category, group)

heatmap_final$data_type <- factor(heatmap_final$data_type,
                                   levels = c("data_null" , "data_varf",
                                              "data_varx","data_varall"))

heatmap_final$category <- reorder(heatmap_final$category, heatmap_final$data_type)

ggplot(heatmap_final,
       aes(x= reorder(value, data_type), y = as.factor(D))) +
  geom_tile(aes(fill = sim_data_quantile)) +
  scale_fill_distiller(palette = "YlGnBu", direction = 1)+ 
  facet_wrap(~category, ncol = 1) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Deciles") +
  xlab("data sets") +
  theme(legend.position = "bottom")+
  scale_x_discrete(breaks = c("data_null" , "data_varf",
                                              "data_varx","data_varall"))

# Clustering all 350 customers
